---
title: "Plan of Attack"
format: pdf
---

```{r}
#| echo: false
#| message: false
#| warning: false
library(tidyverse)
library(gt)

# Load data
data <- read_csv('../data/hillstrom_data.csv')
```


**Research question**: Does the causal effect of email marketing on revenue depend on how recently a customer last purchased, and what does this imply about how retailers should allocate promotional campaigns?

Email campaigns are not free — there is a cost to sending, and not every customer responds. A natural hypothesis is that a customer who purchased two months ago is easier to re-engage than one who has not purchased in nearly a year. If email effectiveness varies systematically with recency, a retailer can do better than blasting the same campaign to everyone: it can concentrate sends where the incremental return is highest and suppress sends where the lift does not cover the cost.

This motivates a more specific operational question: for each recency based customer group (Warm, Cooling, Dormant, Very dormant), should the retailer send the Mens E-Mail campaign, the Womens E-Mail campaign, or no email, in order to maximize incremental profit at an assumed $0.01 per email send cost?


## Section 1: Raw Daw Description

The Hillstrom Email Marketing Experiment dataset contains 64,000 observations. Each observation is uniquely identified by a customer (each row is one customer; no duplicates). Each row represents a single customer, with pre-experiment attributes and post-experiment outcomes. The data covers a single two-week post-campaign window. Key fields are:

```{r}
#| echo: false
tibble(
  Field = c(
    "`recency`", "`history`", "`history_segment`",
    "`mens`", "`womens`", "`zip_code`",
    "`newbie`", "`channel`", "`segment`",
    "`visit`", "`conversion`", "`spend`"
  ),
  Description = c(
    "Months since last purchase",
    "Dollar value of past purchases",
    "Binned dollar value of past purchases e.g. $100 - $200",
    "1 if customer purchased men's merchandise in the past",
    "1 if customer purchased women's merchandise in the past",
    "Urban, suburban, or rural",
    "1 if new customer in the past 12 months",
    "Purchase channel: phone, web, or multichannel",
    "Treatment group: Mens E-Mail, Womens E-Mail, or No E-Mail",
    "1 if the customer visited the website in the following 2 weeks",
    "1 if the customer made a purchase",
    "Dollar amount spent"
  )
) |>
  gt() |>
  fmt_markdown(columns = Field)
```

The data covers 64,000 customers randomly assigned in equal thirds to the three groups (approximately 21,333 per arm). Our dependent variables are `visit`, `conversion`, and `spend`. Our key independent variable is `segment` (treatment group). Pre-experiment covariates (`history`, `history_segment`, `recency`, `mens`, `womens`, `zip_code`, `newbie`, `channel`) are available for heterogeneous-effect exploration.

Summary statistics for the key variables are in the table below. The three dependent variables are: `visit`, `conversion`, and `spend`. The treatment variable (indicator) is `segment`. The remaining fields are pre-treatment covariates used for balance checks and
heterogeneous-effects analysis.


```{r}
#| echo: false
#| message: false
#| warning: false

summary_table <- data %>%
  mutate(
    `Mens E-Mail (0/1)` = if_else(segment == "Mens E-Mail", 1, 0),
    `Womens E-Mail (0/1)` = if_else(segment == "Womens E-Mail", 1, 0),
    `No E-Mail (0/1)` = if_else(segment == "No E-Mail", 1, 0)
  ) %>%
  select(
    `visit (0/1)` = visit,
    `conversion (0/1)` = conversion,
    `spend ($)` = spend,
    `Mens E-Mail (0/1)`,
    `Womens E-Mail (0/1)`,
    `No E-Mail (0/1)`,
    `history ($)` = history,
    `recency (months)` = recency,
    `mens (0/1)` = mens,
    `womens (0/1)` = womens,
    `newbie (0/1)` = newbie
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "value"
  ) %>%
  group_by(Variable) %>%
  summarize(
    Mean = mean(value, na.rm = TRUE),
    SD = sd(value, na.rm = TRUE),
    Min = min(value, na.rm = TRUE),
    Median = median(value, na.rm = TRUE),
    Max = max(value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    across(
      c(Mean, SD, Min, Median, Max),
      ~ round(.x, 3)
    )
  )

summary_table |>
  gt(rowname_col = "Variable") |>
  tab_caption("Summary statistics, full sample (64,000 customers)") |>
  tab_row_group(label = "Dependent variables",      rows = 1:3) |>
  tab_row_group(label = "Treatment (segment)",      rows = 4:6) |>
  tab_row_group(label = "Pre-treatment covariates", rows = 7:11)
```


**Contingencies**: 

## Section 2: Sample Construction Plan

1. **Sample window.** The experiment ran at a single point in time, so there is no sample window decision. Use all 64,000 observations because the dataset itself is the entire experiment. There is no sub-period to carve out.

2. **Sample restrictions.** No observations will be dropped for the main intent-to-treat analysis. Dropping observations post-randomization breaks the random-assignment guarantee and biases the estimate. For secondary heterogeneous-effect exploration, we may subset to men (`mens == 1`) when evaluating the Mens E-Mail treatment and to `women` when evaluating the Womens E-Mail treatment, but those are supplementary analyses, not the main specification. All 64,000 observations are retained for the headline estimate.

3. **Dependent variable integrity.** 
- `visit` and `conversion`: Binary; check coding is $\set{0 ,1}$ with no missing values.
- `spend`: Continuous and right-skewed. For the headline estimate, use ther aw mean difference between arms. For sensitivity, winsorize at the 99th percentile and repor the mean difference again. 

A few big spenders can move the raw mean a lot. The winsorized version shows whether the headline is driven by those outliers or by the bulk of customers.

4. **Independent variables.** `segment` is already clean (3 levels). We construct two specifications:
- `email_any`: 1 if segment is Mens E-Mail or Womens E-Mail, 0 if No E-Mail. Used for the headline "any email" effect.
- `email_mens`, `email_womens`: Separate dummies for each treatment arm, for comparing the two campaigns separately.

The three-arm design lets us test both overall email lift (email vs. no-email) and whether gender-targeted campaigns differ. Constructing both specifications covers both questions without committing prematurely to one framing. No merging or aggregation is needed; all fields are in the single customer-level file.

## Section 3: Analysis Plan

1. **Research design.** Our question is causal: Which pre-campaign customer segments should receive Men’s Email, Women’s Email, or no email to maximize incremental two-week revenue? Should the retailer send promotional emails broadly, or should it target only the customer segments where the randomized experiment shows the largest causal lift in spending? Which customer segments, if any, show enough incremental revenue from Men’s or Women’s promotional emails to justify targeted future campaigns?


2. **Key analyses.** 

